---
title: "sandbox-02-missing-data"
author: "Joshua Rosenberg"
date: "12/29/2017"
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, cache = F, echo = F)
```

```{r, loading-packages}
library(tidyverse)
library(lmerTest)
library(lme4)
library(corrr)
library(jmRtools)
library(tidyttest)
```

```{r, loading-data, eval = F}
esm <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-esm.csv")
pre_survey_data_processed <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-pre-survey.csv")
post_survey_data_partially_processed <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-post-survey.csv")
video <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-video.csv")
pqa <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-pqa.csv")
attendance <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-attendance.csv")
class_data <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-class-video.csv")
demographics <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-demographics.csv")
pm <- read_csv("/Volumes/SCHMIDTLAB/PSE/Data/STEM-IE/STEM-IE-program-match.csv")
```

```{r}
# save.image("~/desktop/sandbox-01.Rdata")
load("~/desktop/sandbox-01.Rdata")
```

```{r, processing-some-data}
attendance <- rename(attendance, participant_ID = ParticipantID)
attendance <- mutate(attendance, prop_attend = DaysAttended / DaysScheduled, 
                     participant_ID = as.integer(participant_ID))
attendance <- select(attendance, participant_ID, prop_attend)

demographics <- filter(demographics, participant_ID!= 7187)
demographics <- left_join(demographics, attendance)

esm$overall_engagement <- jmRtools::composite_mean_maker(esm, hard_working, concentrating, enjoy, interest)
```

```{r, joining-df}
df <- left_join(esm, pre_survey_data_processed, by = "participant_ID") # df & post-survey
df <- left_join(df, video, by = c("program_ID", "response_date", "sociedad_class", "signal_number")) # df & video
df <- left_join(df, demographics, by = c("participant_ID", "program_ID")) # df and demographics
```

```{r, proc-variables, echo = F}
df$participant_ID <- as.factor(df$participant_ID)
df$program_ID <- as.factor(df$program_ID)
df$beep_ID <- as.factor(df$beep_ID)
df$beep_ID_new <- as.factor(df$beep_ID_new)

df$youth_activity_rc <- ifelse(df$youth_activity == "Off Task", "Not Focused", df$youth_activity)

df$youth_activity_rc <- ifelse(df$youth_activity_rc == "Student Presentation" | df$youth_activity_rc == "Problem Solving", "Creating Product", df$youth_activity_rc)

df$youth_activity_rc <- ifelse(df$youth_activity_rc == "Showing Video", "Program Staff Led", df$youth_activity_rc)

df$youth_activity_rc <- as.factor(df$youth_activity_rc)

df$youth_activity_rc <- forcats::fct_relevel(df$youth_activity_rc, "Not Focused")

df$relevance <- jmRtools::composite_mean_maker(df, use_outside, future_goals, important)
```

```{r processing-demographics, echo = F, eval = T}
df$urm <- ifelse(df$race %in% c("White", "Asian"), 0, 1)
df$race <- as.factor(df$race)
df$race <- fct_lump(df$race, n = 2)
df$race_other <- fct_relevel(df$race, "Other")
df$gender_female <- as.factor(df$gender) # female is comparison_group
df$gender_female <- ifelse(df$gender_female == "F", 1, 
                           ifelse(df$gender_female == "M", 0, NA))
```

```{r}
x <- df %>% 
    count(program_ID, sociedad_class, response_date, signal_number) %>% arrange(program_ID, response_date) %>% 
    count(program_ID, sociedad_class) %>% 
    rename(number_of_possible_responses = nn)

xx <- df %>% 
    group_by(program_ID, sociedad_class) %>% 
    count(participant_ID) %>% 
    rename(number_of_responses = n) %>% 
    ungroup()

d <- left_join(xx, x, by = c("program_ID", "sociedad_class")) %>% 
    group_by(participant_ID) %>% 
    summarize(number_of_possible_responses = sum(number_of_possible_responses),
              number_of_responses = sum(number_of_responses)) %>% 
    mutate(response_rate = number_of_responses / number_of_possible_responses) %>%
    write_csv("individual-level-response-rate.csv")
# ps <- read_csv("/Users/joshuarosenberg/Google Drive/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-pre-survey.csv")
# ps
# dd <- left_join(d, ps, by = "participant_ID")
```

# Differences in response rate by demographic and other characteristics

In this document, we examine differences in response rate for STEM-IE by:

* Race
* URM status
* Initial (pre) interest
* Gender
´
We use ANOVA and t-test for group mean differences, and correlation for differences by pre-interest.

We find that there are URM and gender differences in response rate, which appear to be substantively important, but that there are not statistically significant differences in URM status or gender.

## Race

There are not statistically significant differences by race. Other includes white (n = 13), multiracial (n = 5), and Asian (n = 14).

```{r}
left_join(df, d, by = "participant_ID") %>% 
    distinct(urm, race, race_other, gender, gender_female, overall_pre_interest, response_rate) %>% 
    group_by(race) %>% 
    summarize(mean_response_rate = mean(response_rate))

ddd <- left_join(df, d, by = "participant_ID") %>% 
    distinct(urm, race, race_other, gender, gender_female, overall_pre_interest, response_rate)

summary(aov(response_rate ~ race, data = ddd))
```

## URM

Students who are in a URM group have a response rate of 60.6% compared to a response rate of 74.1% for non-URM students (t = 2.622, p = .013, d = .56).

```{r}
left_join(df, d, by = "participant_ID") %>% 
    distinct(urm, race, race_other, gender, gender_female, overall_pre_interest, response_rate) %>% 
    group_by(urm) %>% 
    summarize(mean_response_rate = mean(response_rate))

left_join(df, d, by = "participant_ID") %>% 
    distinct(urm, race, race_other, gender, gender_female, overall_pre_interest, response_rate) %>% 
    t_test(response_rate, urm)

ddd <- left_join(df, d, by = "participant_ID") %>% 
    distinct(urm, race, race_other, gender, gender_female, overall_pre_interest, response_rate)

t.test(ddd$response_rate ~ ddd$urm)
```

## Gender

Female's response rate is 66.2% compared to male's, which is 58.4%; this is a significant difference (t = 2.093, p = .038, d = .31)

```{r}
left_join(df, d, by = "participant_ID") %>% 
    distinct(urm, race, race_other, gender, gender_female, overall_pre_interest, response_rate) %>% 
    group_by(gender) %>% 
    summarize(mean_response_rate = mean(response_rate))

left_join(df, d, by = "participant_ID") %>% 
    distinct(urm, race, race_other, gender, gender_female, overall_pre_interest, response_rate) %>% 
    t_test(response_rate, gender)

ddd <- left_join(df, d, by = "participant_ID") %>% 
    distinct(urm, race, race_other, gender, gender_female, overall_pre_interest, response_rate)

t.test(ddd$response_rate ~ ddd$gender)
```

## Pre-interest

Overall pre-interest and response rate are not correlated.

```{r}
ddd <- left_join(df, d, by = "participant_ID") %>% 
    distinct(urm, race, race_other, gender, gender_female, overall_pre_interest, response_rate) %>% 
    select(overall_pre_interest, response_rate)

cor.test(ddd$overall_pre_interest, ddd$response_rate)
```