# Results for Research Question #1

```{r, setup-results-1, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE, 
                      collapse = TRUE,
                      error = TRUE,
                      fig.width = 6,
                      fig.asp = .618,
                      out.width = "80%", 
                      fig.align = "center", 
                      results = "hold",
                      knitr.kable.na = '') 

```

## Frequency of work with data

From the coding with the STEM-PQA, work with data appears common. Out of the 248 segments, 236 were coded for instructional support for work with data; for the other, not-coded segments, issues with the video-recordings were the primary source of the missing data; in these cases, students may have still replied to signals, but it was not possible to code for instructional support for work with data associated with these responses. 

```{r}
data_frame(Aspect = c("Asking Questions", "Making Observations", "Generating Data", "Data Modeling", "Communicating Findings"),
           Proportion = c(.389, .258, .453, .288, .470)) %>% 
    knitr::kable(booktabs = TRUE, linesep = "", caption = "Proportion of signals for which each of the aspects of work with data was present")
```
