## Results for Research Question #4: How do youth characteristics relate to their engagement in summer STEM programs?

```{r, loading-spreadsheets, eval = FALSE}

library(googlesheets)
library(dplyr)

g1 <- gs_title("USE THIS! New Coding Frame - KMS")
g2 <- gs_title("USE THIS! New Coding Frame - HM")
g3 <- gs_title("USE THIS! New Coding Frame - KS")

d1 <- gs_read(g1, ws = 1)
d2 <- gs_read(g2, ws = 1)
d3 <- gs_read(g3, ws = 1)

d1 <- dplyr::select(d1, program_name, response_date, signal_number, KMS_qual = `Qualitative Coding`)
d2 <- dplyr::select(d2, program_name, response_date, signal_number, HM_qual = `Qualitative Coding`)
d3 <- rename(d3, KLS_qual = `Qualitative Coding`)

d3 <- d3 %>%
  left_join(d2) %>%
  left_join(d1)

d_proc <- d3 %>%
  select(everything(), contains("qual"), -`Josh notes`, -Initials)

# d1i <- dplyr::pull(d1, `Qualitative Coding`)[!d1_na]
# d2i <- dplyr::pull(d2, `Qualitative Coding`)[!d2_na]
# d3i <- dplyr::pull(d3, `Qualitative Coding`)[!d3_na]

readr::write_csv(d_proc, "qual-coding.csv")
```

```{r, read-proc-sheet, eval = F}
qual_sheets <- readr::read_csv("qual-coding.csv")
```

### Summary of Findings for Research Question #4

These findings focused on the affordances and constrained of work with data in summer STEM programs and how instructional support for work with data was measured. Broadly, the qualitative analysis suggested possible explanations for the findings for research questions #2 and #3. For these questions, little variability was found to exist at the momentary level, and the predictors at the momentary level (instructional support for work with data) and at the youth level (pre-interest, gender, and URM status) demonstrated modest relations with the profiles. These relations can be due to a variety of reasons, particularly 1) how the variables for the PECs and how instructional support for work with data is measured, and 2) how suitable of summer STEM programs for work with data. Accordingly, this analysis resulted in findings organized around the following two themes. The first theme concerned *affordances and constraints of summer STEM programs for work with data*. The second concerned *howinstructional support for work with data was measured*. Both are described in the remainder of this section. Another possible explanation related to whether PECs and the variables that make them up are appropriate outcomes, and how the PECS are measured, is an important question, but one that cannot readily be assessed from the video data that was analyzed; however, this topic is explored in the next chapter.
